### 2018-01-14 Atomic Spectra/paper.tex
fixed that table

\input{header}
\physics
\begin{document}
\papertitle{Analysis of Atomic Spectra of Various Gases}
\paperauth{A}{Khesin}{Lab Technician}
\paperauth{P}{Zavyalova}{Data Analyst}
\paperdate{January 15, 2018}
\begin{paperabs}
Abstract goes here
\end{paperabs}

\begin{paper}
\papersec{Introduction}

In this exercise, a prism spectrometer is calibrated and used to determine the Rydberg constant \( R_H \). Further, an ``unknown'' gas is identified by measuring its spectral lines and comparing them with the obtained calibration curve.

\papersec{Method}

Method goes here

\papersec{Data}

(Helium)

\begin{tabular}{|l|c|r|}
	\hline
	Vernier & Wavelength (nm) & Color \\ \hline
	6.66 & 706.5 & red \\ \hline
	7.06 & 667.8 & red \\ \hline
	8.31 & 587.6 & yellow \\ \hline
	11.09 & 501.6 & green \\ \hline
	11.50 & 492.2 & green \\ \hline
	12.58 & 471.3 & blue \\ \hline
	14.19 & 447.1 & blue \\ \hline
\end{tabular}

\noindent\begin{papertable}{|[outer]I|[inner]C|[inner]C|[inner]C|[outer]}
\paperoline\papertableindexheader&\textsc{Scale}&\textsc{Wavelength}&\multirow{2}{*}{\textsc{Colour}}\\
\papertablecheaderunit{.02}{\hspace{-0.6ex}}\papertablecheadersymbol{\lambda~(\si{\nano\meter})}&\\\paperiline
\papertableindex\papertablecval{6.66}\papertablecval{706.5}\papertablecval{Red}\\\paperiline
\papertableindex\papertablecval{7.06}\papertablecval{667.8}\papertablecval{Red}\\\paperiline
\papertableindex\papertablecval{8.31}\papertablecval{587.6}\papertablecval{Yellow}\\\paperiline
\papertableindex\papertablecval{11.09}\papertablecval{501.6}\papertablecval{Green}\\\paperiline
\papertableindex\papertablecval{11.50}\papertablecval{492.2}\papertablecval{Green}\\\paperiline
\papertableindex\papertablecval{12.58}\papertablecval{471.3}\papertablecval{Blue}\\\paperiline
\papertableindex\papertablecval{14.19}\papertablecval{447.1}\papertablecval{Blue}\\\paperoline
\end{papertable}

(Hydrogen)

\begin{tabular}{|l|r|}
	\hline
	Vernier & Color \\ \hline
	7.26 & red \\ \hline
	11.80 & blue-green \\ \hline
	15.15 & violet \\ \hline
	17.75 & violet \\ \hline
\end{tabular}

(Unknown \# 7)

\begin{tabular}{|l|r|}
	\hline
	Vernier & Color \\ \hline
	8.44 & orange \\ \hline
	9.18 & green \\ \hline
	13.89 & violet \\ \hline
	15.38 & violet \\ \hline
\end{tabular}

\papersec{Analysis}

Analysis goes here

\papersec{Conclusion}

Conclusion goes here
\end{paper}
\end{document}