### 2018-01-17 lenses/paper.tex
Added lenses setup pdf

\input{header}
\physics
\begin{document}
\papertitle{Thick Lenses}
\paperauth{A}{Khesin}{1002442029}
\paperauth{P}{Zavyalova}{1002345036}
\paperdate{January 22, 2018}
\begin{paperabs}

	Abstract

\end{paperabs}

\begin{paper}

\papersec{Introduction}

	Introduction

\papersec{Observations}

	Observations

\papersec{Analysis}

	Analysis

\papersec{Conclusion}

		Conclusion

\papersec{Sources}

	Sources

\end{paper}

\end{document}