### 2018-01-29 circuit-elements/paper.tex
Added polarization data plots

\input{header}
\physics
\begin{document}
\papertitle{Nonlinear Circuit Elements}
\paperauth{A}{Khesin}{1002442029}
\paperauth{P}{Zavyalova}{1002345036}
\paperdate{January 17, 2018}
\begin{paperabs}
	The relationships between current, potential, resistance, and temperature are of particular interest in nonlinear circuit components. In this experiment, these physical quantities were investigated for a thermistor and an incandescent light bulb. For the thermistor, resistance was found to decrease exponentially as temperature increased, following the relation \( R = \pars{1.0 \pm .1} \times 10^{13}e^{-T/\pars{24.7 \pm .3}} \). The relationship between potential and current of the light bulb was determined to obey a power law, \( V = \pars{0.067 \pm .002} I^{\pars{1.63 \pm .01}} \). Further, the exponential relational constant in Stefan-Boltzmann law was found to be $\pars{4.17\pm.05}$.

\end{paperabs}

\begin{paper}

\papersec{Introduction}

	Circuit components may be classified based on how they process signals as either linear or nonlinear. In linear components, such as resistors, capacitors, and some inductors, the signal obeys the principle of superposition; such components process electrical signals without distorting them. Nonlinear circuit elements do not have a linear relationship between current and potential, and thus may distort electrical signals.

	The purpose of this experiment was to study the relationship between voltage and current for a thermistor and an incandescent lamp. These circuit elements are nonlinear; for a thermistor, resistance is a function of temperature and obeys the following relation.

	\papereq{thermistor_model}{R = R_0e^{ -\frac{T}{T_0} } .}{Serbanescu}
	\begin{paperwhere}
		\papervar{T}{absolute temperature}{\kelvin}
		\papervar{T_0}{constant of device}{\kelvin}
		\papervar{R_0}{maximum resistance (device specific)}{\ohm}
	\end{paperwhere}

	As evident from the above expression, the thermistor's resistance decreases with increasing temperature. This is unlike resistance of metals (such as the filament of a light bulb examined further in the experiment), which exhibit the opposite relationship.

	The relation in (1) was investigated using a multimeter, a thermistor, a thermometer and a thermal flask for maintaining the thermistor at different temperatures, set up as shown below. Temperature of the thermistor was decreased by adding cold water to the thermal flask and stirring to ensure it was uniform. \vspace{-0.8em}

	\paperfig{setup1}{\pdf{setup1} \vspace{-2.5em} }{The experimental setup used to measure resistance of a thermistor at varying temperatures. \vspace{0.5em}}

	In general, the tungsten filament in an incandescent light bulb obeys Ohm's law. As the temperature increases, however, the resistance of the metal increases and above a certain temperature becomes proportional to the absolute temperature. The actual relationship between \( V \) and \( I \) was investigated in this experiment by fitting obtained measurements of current and potential to the following.

	\papereq{bulb-model}{ V = k I^a .}{Serbanescu}
	\begin{paperwhere}
		\papervar{V}{potential}{\volt}
		\papervar{I}{current}{\ampere}
		\papervar{k}{proportionality constant}{}
		\papervar{a}{exponential relational constant}{}
	\end{paperwhere}

	Potential and current were measured using two multimeters, with one measuring potential and the other current, a power supply, a tungsten filament light bulb, and a switch arranged as shown.

	\paperfig{setup2}{\pdf{setup2} \vspace{-2.5em}}{Experimental setup used to obtain potential and current data for the light bulb. \vspace{0.5em}}

	The current-potential data obtained for an incandescent light bulb was further used to determine the value of \( n \) in the Stefan-Boltzmann law.

	\papereq{bulb-model}{ \kappa = \epsilon \sigma T^n. }{Serbanescu}
	\begin{paperwhere}
		\papervar{\kappa}{power per unit area}{\watt\per\meter\squared}
		\papervar{\epsilon}{emissivity of radiating body}{}
		\papervar{\sigma}{Stefan-Boltzmann constant}{\watt\per\meter\squared\per\kelvin\tothe{4}}
		\papervar{T}{absolute temperature}{K}
		\papervar{n}{exponential relational constant}{}
	\end{paperwhere}

	Using the expression in (3) and assuming thermal equilibrium, \( \kappa \) and \( T \) were expressed in terms of \( I \) and \( V \). Additionally, by expressing \( V \) in terms of \( I \) as shown in (2), \( \kappa \) was written in terms of \( I \) and constants only.

	\papereq{kappa}{ \kappa = \frac{VI}{A} = \frac{k}{A} I^{(a+1)}. }{}
	\begin{paperwhere}
		\papervar{A}{surface area of filament}{\si{\meter\squared}}
	\end{paperwhere}

	By using the fact that resistance of a pure metal is directly proportional to its temperature (\( R = k'T \), with \( k' \) being a constant of proportionality) it was shown that

	\papereq{temperature}{ T = \frac{1}{k'} \frac{V}{I} =  \frac{k}{k'} I^{(a-1)}.}{}

	Finally, combining the above expressions and setting constants to be equal, the following relationship for \( n \) in terms of \( a \) was obtained. \vspace{-0.5em}

	\papereq{temperature}{ n = \frac{a+1}{a-1} }{}

\papersec{Observations}

	The resistance of the thermistor at various temperatures was measured. The uncertainty, as specified by the manufacturer, was $ \pm (.2\% + 1 $ LSD) for resistances in the $ (2 \si{\kilo\ohm}, 200 \si{\kilo\ohm}) $ range. The reading error of the thermometer was taken to be $ \pm .5 \si{\kelvin} $. \vspace{0.5em}

	\paperfig{thermistor}{\begin{papertable}{|[outer]I|[inner]C|[inner]U|[outer]}
			\paperoline\papertableindexheader&\textsc{Temperature}&\multicolumn{2}{c|[outer]}{\textsc{Resistance}}\\
			\papertablecheadersymbol{T$ $(\pm.5$ $\si{\kelvin})}\papertableuheadersymbole{R$ $(\si{\kilo\ohm})}\\\paperiline
			\papertableindex\papertablecval{352.5}\papertableuval{6.68}{.02}\\\paperiline
			\papertableindex\papertablecval{349.0}\papertableuval{6.93}{.02}\\\paperiline
			\papertableindex\papertablecval{348.0}\papertableuval{7.54}{.03}\\\paperiline
			\papertableindex\papertablecval{346.0}\papertableuval{8.01}{.03}\\\paperiline
			\papertableindex\papertablecval{340.0}\papertableuval{9.95}{.03}\\\paperiline
			\papertableindex\papertablecval{338.0}\papertableuval{10.33}{.03}\\\paperiline
			\papertableindex\papertablecval{335.0}\papertableuval{11.76}{.03}\\\paperiline
			\papertableindex\papertablecval{334.5}\papertableuval{12.36}{.03}\\\paperiline
			\papertableindex\papertablecval{329.5}\papertableuval{14.45}{.04}\\\paperiline
			\papertableindex\papertablecval{327.5}\papertableuval{15.33}{.04}\\\paperiline
			\papertableindex\papertablecval{324.0}\papertableuval{17.23}{.04}\\\paperiline
			\papertableindex\papertablecval{321.0}\papertableuval{19.51}{.05}\\\paperiline
			\papertableindex\papertablecval{319.0}\papertableuval{21.7}{.1}\\\paperiline
			\papertableindex\papertablecval{317.0}\papertableuval{24.3}{.1}\\\paperiline
			\papertableindex\papertablecval{314.0}\papertableuval{26.1}{.2}\\\paperiline
			\papertableindex\papertablecval{309.5}\papertableuval{31.4}{.2}\\\paperiline
			\papertableindex\papertablecval{308.5}\papertableuval{33.7}{.2}\\\paperiline
			\papertableindex\papertablecval{307.5}\papertableuval{35.5}{.2}\\\paperiline
			\papertableindex\papertablecval{305.0}\papertableuval{38.0}{.2}\\\paperiline
			\papertableindex\papertablecval{304.0}\papertableuval{40.3}{.2}\\\paperiline
			\papertableindex\papertablecval{301.5}\papertableuval{44.3}{.2}\\\paperiline
			\papertableindex\papertablecval{299.5}\papertableuval{47.7}{.2}\\\paperiline
			\papertableindex\papertablecval{298.0}\papertableuval{51.6}{.2}\\\paperiline
			\papertableindex\papertablecval{296.0}\papertableuval{56.3}{.2}\\\paperiline
			\papertableindex\papertablecval{294.0}\papertableuval{61.3}{.2}\\\paperiline
			\papertableindex\papertablecval{292.5}\papertableuval{66.9}{.2}\\\paperiline
			\papertableindex\papertablecval{290.5}\papertableuval{73.8}{.2}\\\paperiline
			\papertableindex\papertablecval{286.0}\papertableuval{92.9}{.3}\\\paperiline
			\papertableindex\papertablecval{283.0}\papertableuval{104.7}{.3}\\\paperiline
			\papertableindex\papertablecval{278.0}\papertableuval{~126.2}{\parbox{\widthof{126.2~}}{.4}}\\\paperoline
		\end{papertable}\vspace{-1.5em}}
	{Thermistor's resistance at given temperatures. }\vspace{0.5em}

	Further, current and potential for a tungsten filament light bulb were measured and recorded. As specified by the manufacturer, the uncertainty in potential measurements was taken to be \( \pm (.25\% + 1 \) LSD); current measurements carried an uncertainty of \( \pm (1\% + 3 \) LSD). \vspace{0.5em}

	\paperfig{light-bulb}{\begin{papertable}{|[outer]I|[inner]U|[inner]U|[outer]}
			\paperoline\papertableindexheader&\multicolumn{2}{c|[inner]}{\textsc{Potential}}&\multicolumn{2}{c|[outer]}{\textsc{Current}}\\
			\papertableuheadersymbol{V$ $(\si{\volt})}\papertableuheadersymbole{I$ $(\si{\milli\ampere)}}\\\paperiline
			\papertableindex\papertableuval{0.0222}{.0002}\papertableuval{0.324}{.003}\\\paperiline
			\papertableindex\papertableuval{0.0776}{.0003}\papertableuval{1.045}{.009}\\\paperiline
			\papertableindex\papertableuval{0.1122}{.0004}\papertableuval{1.39}{.01}\\\paperiline
			\papertableindex\papertableuval{0.244}{.002}\papertableuval{2.27}{.03}\\\paperiline
			\papertableindex\papertableuval{0.472}{.002}\papertableuval{3.40}{.04}\\\paperiline
			\papertableindex\papertableuval{0.541}{.002}\papertableuval{3.69}{.04}\\\paperiline
			\papertableindex\papertableuval{0.697}{.003}\papertableuval{4.28}{.04}\\\paperiline
			\papertableindex\papertableuval{0.871}{.003}\papertableuval{4.89}{.05}\\\paperiline
			\papertableindex\papertableuval{1.128}{.004}\papertableuval{5.72}{.05}\\\paperiline
			\papertableindex\papertableuval{1.840}{.006}\papertableuval{7.68}{.07}\\\paperiline
			\papertableindex\papertableuval{1.940}{.006}\papertableuval{7.93}{.07}\\\paperiline
			\papertableindex\papertableuval{4.01}{.02}\papertableuval{12.3}{.1}\\\paperiline
			\papertableindex\papertableuval{6.00}{.03}\papertableuval{15.8}{.1}\\\paperiline
			\papertableindex\papertableuval{8.01}{.03}\papertableuval{18.8}{.2}\\\paperiline
			\papertableindex\papertableuval{10.24}{.04}\papertableuval{21.8}{.3}\\\paperiline
			\papertableindex\papertableuval{12.07}{.04}\papertableuval{24.1}{.3}\\\paperiline
			\papertableindex\papertableuval{14.27}{.05}\papertableuval{26.6}{.3}\\\paperiline
			\papertableindex\papertableuval{16.01}{.05}\papertableuval{28.5}{.3}\\\paperiline
			\papertableindex\papertableuval{17.98}{.05}\papertableuval{30.6}{.3}\\\paperiline
			\papertableindex\papertableuval{18.45}{.06}\papertableuval{31.1}{.3}\\\paperiline
			\papertableindex\papertableuval{19.86}{.06}\papertableuval{32.5}{.3}\\\paperiline
			\papertableindex\papertableuval{21.0}{.2}\papertableuval{33.6}{.4}\\\paperiline
			\papertableindex\papertableuval{22.1}{.2}\papertableuval{34.6}{.4}\\\paperiline
			\papertableindex\papertableuval{22.9}{.2}\papertableuval{35.4}{.4}\\\paperiline
			\papertableindex\papertableuval{24.0}{.2}\papertableuval{36.4}{.4}\\\paperiline
			\papertableindex\papertableuval{25.1}{.2}\papertableuval{37.3}{.4}\\\paperoline
			\end{papertable}\vspace{-1.5em}}
	{Potential and current through the light bulb.}

\papersec{Analysis}

	Upon propagating uncertainties in the obtained data using the method of quadratures, the relationship between the temperature and resistance of a thermistor was calculated using the expression in (1) as the model function. Least squares fitting of all data points yielded \( R_0 = (1.0 \pm .1) \times 10^{13} \si{\ohm} \) and \( T_0 = (24.7 \pm .3) \si{\kelvin} \) as the fit parameters, with a reduced \( \chi ^ 2 \) value of \( 252.1710 \). The data, fit, and residuals are shown below.

	A visualization of residuals  indicated that the data can be better modeled at higher temperatures. Least squares fitting without the data points corresponding to three lowest temperatures yielded a reduced \( \chi ^ 2 \) value of \( 181.5543 \). Removing more lower-temperature points improved it further.

	\paperfig{thermistor-plot}{\pdf{Thermistor}}{Resistance and residuals plotted against the temperature in the thermal flask. The curve has the function \( (1.0 \pm .1) \times 10^{13}e^{T/\pars{24.7 \pm .3}} \) and an \( R^2 \) value of 0.992783. Error bars on both temperature and resistance were included but were too small to be visible.}

	Further, a least squares fitting was performed on the data obtained for the incandescent light bulb, with the expression in (2) used as a model function. The formula was found to hold; the resulting parameters were \( k =  (0.067 \pm .002) \si{\volt\per\ampere} \) and \( a = (1.63 \pm .01) \). The value of reduced \( \chi ^ 2 \) obtained was \( 176.8952 \). The data, model, and residuals for this fit are shown below.

	\paperfig{bulb-plot}{\pdf{Bulb}}{Potential and residuals plotted against current through the incandescent bulb. The curve has the function \( \pars{0.067 \pm .002} I^{\pars{1.63 \pm .01}} \) with an \( R^2 \) value of \( 0.999532 \). Error bars on the data points were included but were too small to be visible.}

	As evidenced above, thermistors and metal resistors have different variations of resistance with temperatures. For metal resistors, resistance increasing with temperature is intuitively explained by considering increasing vibrations of electrons in the metal. At higher temperatures, electrons will interfere more with free currents and thus the resistance will increase. Thermistors, on the other hand, are semiconductors. Increasing the temperature increases the number of available electrons, the conductivity increases, and thus the resistance decreases.

	Upon obtaining the value of \( a \), the universal constant \( n \) in Stefan-Boltzmmann law could be approximated. Plugging \( a = (1.63 \pm .01) \) into (6) yielded $\pars{4.17\pm.05}$ (uncertainty propagated using the method  of quadratures), which was close the actual value of \( n = 4\). Another constant appearing in Stefan-Boltzmann law, \( \epsilon \) (emissivity of the radiating body), differs depending on the material. \( \epsilon = 1 \) corresponds to a body that is a perfect absorber and emitter of radiant energy (i.e., a black body); \( \epsilon = 0 \) corresponds to a perfect reflector of radiant energy.

\papersec{Conclusion}

		This experiment has investigated two nonlinear circuit elements, a thermistor and an incandescent tungsten filament light bulb.\vfill\phantom{.}\columnbreak\noindent Resistance of the thermistor was found to be a function of the temperature, satisfying the exponential relation in (1) with \( R_0 = (1.0 \pm .1) \times 10^{13} \si{\ohm} \) and \( T_0 = (24.7 \pm .3) \si{\kelvin} \).  For the light bulb, the relationship between potential and current was shown to follow a power law introduced in (2), with \( k =  (0.067 \pm .002) \si{\volt\per\ampere} \) and \( a = \pars{1.63 \pm .01} \).

		The most significant sources of error arose from potentially uneven temperature of water in the thermal flask with the thermistor where there could have been momentary hot or cold pockets and inconsistencies in current and potential readings of the multimeter when switching the range of input signal on the multimeters.

		The above value of \( a \) was further used to approximate the power \( n \) of \( T \) in Stefan-Boltzmann law; \( n \) was found to be \( 4.17 \pm .05 \), reasonably close to the actual value of \( n = 4 \).

\papersec{Sources}

	\papersource{Serbanescu, R., Non-linear Circuit Elements, 2015}

\end{paper}

\end{document}