### 2018-03-12 two_terminal/paper.tex
Paper + characteristic curves

\input{header}
\usepackage{tikz}
\usetikzlibrary{optics}

\physics
\begin{document}
\papertitle{Two-Terminal Devices}
\paperauth{A}{Khesin}{1002442029}
\paperauth{P}{Zavyalova}{1002345036}
\paperdate{March 19, 2018, Completed March 13, 2018}
\begin{paperabs}

	Abstract goes here

\end{paperabs}

\begin{paper}

\papersec{Introduction}

	Introduction goes here

\papersec{Observations}

	Observations go here

\papersec{Analysis}

	Analysis goes here

\papersec{Conclusion}

	Conclusion goes here

\papersec{Sources}

	\papersource{J.V., R.M.S., Two-terminal devices, 2010}

\end{paper}
\end{document}