### 2018-03-21 franck-hertz/paper.tex
Franck-Hertz plots

\input{header}
\physics
\begin{document}
\papertitle{The Franck-Hertz Experiment}
\paperauth{A}{Khesin}{1002442029}
\paperauth{P}{Zavyalova}{1002345036}
\paperdate{March 121, 2018, Completed March 15, 2018}
\begin{paperabs}

	Abstract goes here

\end{paperabs}

\begin{paper}

\papersec{Introduction}

	Introduction goes here

\papersec{Observations}

	Observations go here

\papersec{Analysis}

	Analysis goes here

\papersec{Conclusion}

	Conclusion goes here

\papersec{Sources}

	\\papersource{Serbanescu, R.,The Franck-Hertz Experiment, 2017}

\end{paper}
\end{document}